---
title: "Antigonish Current Data"
format: 
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi = 600, fig.width=8)

library(adcp)
library(dplyr)
library(DT)
library(gggibbous) 
library(ggplot2)
library(here)
library(leaflet)
library(lubridate) 
library(patchwork) 
library(purrr)  
library(RColorBrewer)
library(readr) 
library(summaryplots) 
library(tidyr) 
library(viridis)

source(here("functions/subchunkify.R"))
source(here("functions/identify_extreme_speed_events.R"))

dt_options <- list(
  dom = 'ft',
  paging = FALSE,
  searching = FALSE,
  columnDefs = list(list(className = 'dt-center', targets = "_all"))
)

theme_set(theme_light())
pal <- brewer.pal(9, "Blues")[5]

depls <- c("AT001", "AT002", "AT003", "AT004")

# import Antigonish Harbour deployments
dat_all <- adcp_import_data(here("data"), county = "antigonish") %>% 
  filter(deployment_id %in% depls) %>% 
  mutate(sea_water_speed_cm_s = sea_water_speed_m_s * 100) %>%
  select(-c(sea_water_speed_m_s, county, waterbody, bin_depth_below_surface_m))

metadata <- adcp_read_nsdfa_metadata(
  path = here("supp_data/2023-08-14 - NSDFA Tracking Sheet.xlsx")) %>% 
  filter(Depl_ID %in% depls) %>% 
  select(
    Station = Station_Name, 
    Latitude = Depl_Lat, 
    Longitude = Depl_Lon, 
    `Deployment Date` = Depl_Date, 
    `Retrieval Date` = Recv_Date, 
    `Depth (m)` = Inst_Depth, 
    Bin_Size, First_Bin_Range, Current_Ensemble_Interval_s, Current_Averaging_Interval_s, Current_PingsPerEnsemble
  )

```

`r Sys.Date()`

# Deployment Locations

```{r}
leaflet(metadata) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    ~Longitude, ~Latitude, 
    fillColor = "black", 
    fillOpacity = 1, 
    stroke = FALSE, 
    label = ~Station, 
    radius = 4,
    labelOptions = labelOptions(
      noHide = T, textOnly = TRUE, offset = c(0.9, 11), textsize = "14px")
  )
```

# ADCP Deployment Metadata

For all deployments:

- Bin Size: `r unique(metadata$Bin_Size)` m
- First Bin Range: `r unique(metadata$First_Bin_Range)` m
- Ensemble Interval: `r unique(metadata$Current_Ensemble_Interval_s)` seconds
- Averaging Interval: `r unique(metadata$Current_Averaging_Interval_s)` seconds
- Pings per Ensemble: `r unique(metadata$Current_PingsPerEnsemble)`


```{r}
metadata %>% 
  select(
    -c(Bin_Size, First_Bin_Range, 
       Current_Ensemble_Interval_s,
       Current_Averaging_Interval_s, 
       Current_PingsPerEnsemble)
    ) %>% 
  datatable(options = dt_options, rownames = FALSE)
```

# Current Speed Overview

- Current speed over time for each location.
- Each panel is for one bin; the number in the grey bar indicates the bin height above the sea floor (in metres).
- Shallowest bins have fewer good observations because of side lobe interference.

## Current Speed at Depth

::: panel-tabset
```{r}
#| results: "asis"

for(i in seq_along(depls)){
  depl_i <- depls[i]

  dat_i <- dat_all %>%
    filter(deployment_id == depl_i)

  bin_heights <- sort(
    unique(dat_i$bin_height_above_sea_floor_m), decreasing = TRUE
  )

  dat_i <- dat_i %>%
    mutate(
      bin_height_above_sea_floor_m =
        ordered(bin_height_above_sea_floor_m, levels = bin_heights)
    )

  # height of figure j
  n_bins <- length(bin_heights)
  h <- n_bins * 1

  cat('\n###', unique(dat_i$station), '\n')

  p <- ggplot(dat_i, aes(timestamp_utc, sea_water_speed_cm_s)) +
    geom_line(col = pal) +
    facet_wrap(~bin_height_above_sea_floor_m, ncol = 1) +
    scale_x_datetime(date_labels = "%Y-%B-%d")

  subchunkify(p, fig_height = h, fig_width = 8.5)
}
```
:::

## Average Current Speed at Depth

Average current speed (+/- 1 standard deviation) for each depth bin.

::: panel-tabset
```{r}
#| results: "asis"
#| message: false

dat_summary <- dat_all %>% 
  group_by(deployment_id, station, bin_height_above_sea_floor_m) %>% 
  summarise(
    mean_sea_water_speed_cm_s = round(mean(sea_water_speed_cm_s), digits = 1),
    sd_sea_water_speed_cm_s = round(sd(sea_water_speed_cm_s), digits = 1),
    q_95 = quantile(sea_water_speed_cm_s, probs = 0.95)
  )

for(i in seq_along(depls)){
  depl_i <- depls[i]
  
  dat_i <- dat_summary %>% 
    filter(deployment_id == depl_i)
  
  bin_heights <- sort(unique(dat_i$bin_height_above_sea_floor_m))
  
  dat_i <- dat_i %>% 
    mutate(
      bin_height_above_sea_floor_m = 
        ordered(bin_height_above_sea_floor_m, levels = bin_heights)
    )

  # height of figure j
  n_bins <- length(bin_heights)
  h <- n_bins * 0.4

  cat('\n###', unique(dat_i$station), '\n')
  
  p <- ggplot(dat_i,
         aes(mean_sea_water_speed_cm_s, 
             bin_height_above_sea_floor_m, 
             col = mean_sea_water_speed_cm_s)) +
    geom_point(size = 4) +
    geom_errorbar(
      aes(xmin = mean_sea_water_speed_cm_s - sd_sea_water_speed_cm_s, 
          xmax = mean_sea_water_speed_cm_s + sd_sea_water_speed_cm_s), 
      width = 0
    ) +
    scale_x_continuous("sea_water_speed_cm_s") +
    theme(
      text = element_text(size = 12),
      legend.position = "none"
    ) 
  
  subchunkify(p, fig_height = h, fig_width = 8.5)
}
```
:::

# Near-Bottom Current Speed 

"Near-bottom" is the deepest bin at each station.

Percent of observations in 12 speed categories for the deepest depth bin at each station.

::: panel-tabset
```{r}
#| results: "asis"
#| message: false

dat_bottom <- dat_all %>% 
  group_by(deployment_id) %>% 
  mutate(deepest_bin = min(bin_height_above_sea_floor_m)) %>% 
  filter(bin_height_above_sea_floor_m == deepest_bin) %>% 
  select(-deepest_bin) %>% 
  ungroup()

n_ints <- 12
speed_cols <- viridis(n_ints, option = "F", direction = -1)

for(i in seq_along(depls)){
  depl_i <- depls[i]
  
  dat_i <- dat_bottom %>% 
    filter(deployment_id == depl_i)
  
  ints <- adcp_count_obs(dat_i, sea_water_speed_cm_s)
  breaks <- c(ints$lower, max(ints$upper))
  
  cat('\n###', unique(dat_i$station), '\n')
  
  p <- adcp_plot_speed_hist(ints, bar_cols = speed_cols)
  
  subchunkify(p, fig_height = 5.5, fig_width = 8.5)
}
```
:::

## Extreme Current Speed

"Extreme values" are defined here as current speeds greater than the 95 th quantile (e.g., top 5 % of current speed observations at each location). 

```{r}
#| message: false

extreme_current <- dat_bottom %>% 
  left_join(
    dat_summary, 
    by = c("deployment_id", "station", "bin_height_above_sea_floor_m")
  ) %>%
  mutate(extreme = sea_water_speed_cm_s > q_95)

extreme_current %>% 
  filter(extreme) %>% 
  group_by(deployment_id, station, bin_height_above_sea_floor_m) %>% 
  summarise(
    mean_speed_extreme_values_cm_s = round(mean(sea_water_speed_cm_s), digits = 1)
  ) %>% 
  ungroup() %>% 
  select(-deployment_id) %>% 
  left_join(dat_summary, by = c("station", "bin_height_above_sea_floor_m")) %>% 
  select(
    station,
    `bin_height m` = bin_height_above_sea_floor_m,
    `mean_speed cm_s` = mean_sea_water_speed_cm_s,
    q_95,
    `mean_extreme_speed cm_s`= mean_speed_extreme_values_cm_s
  ) %>% 
  mutate(
    speed_ratio = round(
      `mean_extreme_speed cm_s` / `mean_speed cm_s`, digits = 2)
  ) %>% 
  datatable(options = dt_options, rownames = FALSE)

```

</br>

## Near-Bottom Current Speed Over Time

Sea water speed over time. Shaded grey area indicates extreme values.
Circles indicate the phase of the moon (yellow circle indicates full moon; black circle indicates new moon). Data from [Moon Phases 2022 – Lunar Calendar for Antigonish, Nova Scotia, Canada](https://www.timeanddate.com/moon/phases/@5886184?year=2022)

::: panel-tabset
```{r}
#| results: "asis"
#| message: false

# data from: https://www.timeanddate.com/moon/phases/@5886184?year=2022
moon_phase <- read_csv(
  here("supp_data/2022_antigonish_moon_phases.csv"),
  show_col_types = FALSE
) %>% 
  select(-c(Lunation, Duration)) %>% 
  pivot_longer(cols = everything(), names_to = "phase", values_to = "date") %>% 
  mutate(
    date = paste0(date, "-2022"),
    date = as_datetime(date, format = "%d-%b-%Y"),
    percent = case_when(
      phase == "New Moon" ~ 0,
      phase == "First Quarter" ~ 0.5,
      phase ==  "Third Quarter" ~ 0.5,
      phase == "Full Moon" ~ 1
    )
  ) %>% 
  filter(
    date >= min(extreme_current$timestamp_utc),
    date <= max(extreme_current$timestamp_utc)
  )

for(i in seq_along(depls)){
  depl_i <- depls[i]
  
  dat_i <- extreme_current %>% 
    filter(deployment_id == depl_i)
  
  moon_height <- max(dat_i$sea_water_speed_cm_s) - 
    0.1 * max(dat_i$sea_water_speed_cm_s)
  
  cat('\n###', unique(dat_i$station), '\n')
  
  p <- ggplot(dat_i, aes(timestamp_utc, sea_water_speed_cm_s)) +
    annotate(
      "rect", fill = "grey80",
      xmin = as_datetime(-Inf),
      xmax = as_datetime(Inf),
      ymin = unique(dat_i$q_95), ymax = Inf
    ) +
    geom_line(col = pal, linewidth = 0.5) +
    geom_moon(
      data = moon_phase, aes(date, moon_height),
      ratio = 1, size = 6, fill = "black") +
    geom_moon(
      data = moon_phase, aes(date, moon_height, ratio = percent),
      size = 6, right = FALSE, fill = "gold"
    ) 
  
  subchunkify(p, fig_height = 5.5, fig_width = 8.5)
}
  
```
:::

## Extreme Current Speed Events

Top panel: duration of extreme current events (e.g., the amount of time when consecutive observations are > 95 th quantile).

Bottom panel: duration of time between extreme current events.

Note the difference in x-axis scale.

::: panel-tabset
```{r}
#| results: "asis"
#| message: false

speed_events <- extreme_current %>% 
  identify_extreme_speed_events() %>% 
  mutate(
    event_duration_hours = difftime(event_end, event_start, units = "hours"),
    event_duration_hours = as.numeric(event_duration_hours)
  ) %>% 
  group_by(station, deployment_id) %>% 
  mutate(
    time_between_events_days = difftime(lead(event_start), event_end, units = "days"),
    time_between_events_days = as.numeric(time_between_events_days)
  )

for(i in seq_along(depls)){
  depl_i <- depls[i]
  
  dat_i <- speed_events %>% 
    filter(deployment_id == depl_i)
  
  cat('\n###', unique(dat_i$station), '\n')
 
  p1 <- plot_histogram(
    dat_i, hist_col = "event_duration_hours", pal = pal, binwidth = 0.25
  ) +
    scale_x_continuous("Extreme Speed Duration (hours)") +
    scale_y_continuous("Percent of Events")
  
  p2 <- plot_histogram(
    na.omit(dat_i), 
    hist_col = "time_between_events_days", 
    binwidth = 0.25,
    pal = "#AE1759FF"
  ) +
    scale_x_continuous("Time Between Extreme Speed Events (days)") +
    scale_y_continuous("Percent of Events")
  
  p <- p1 / p2
  
  subchunkify(p, fig_height = 8, fig_width = 8.5)
}
```
:::

## Current Speed & Direction Over Time

::: panel-tabset
```{r}
#| results: "asis"
#| message: false

for(i in seq_along(depls)){
  depl_i <- depls[i]
  
  dat_i <- extreme_current %>% 
    filter(deployment_id == depl_i)
  
  cat('\n###', unique(dat_i$station), '\n')
  
  p <- ggplot(
    dat_i, 
    aes(timestamp_utc, sea_water_speed_cm_s, 
        col = sea_water_to_direction_degree)) +
    annotate(
      "rect", fill = "grey80",
      xmin = as_datetime(-Inf),
      xmax = as_datetime(Inf),
      ymin = unique(dat_i$q_95), ymax = Inf
    ) +
    geom_line(linewidth = 0.5) +
    scale_color_viridis("to_direction", direction = -1, option = "C")
  
  subchunkify(p, fig_height = 5, fig_width = 8.5)
}
```

:::
